Nested Stochastic Block Models Applied to the Analysis of Single Cell Data

### 2021-07-15 manuscript.tex
Modified Figure 1

\begin{figure}[H]
\centering
\includegraphics[keepaspectratio,width=0.6\textwidth,height=\textheight]{FIgure_Random.png}
\caption[]{nSBM results for randomised data. (A) UMAP embedding of ~3k PBMC after standard processing with Leiden approach (left) or nSBM (right). Cell grouping is consistent for both the approaches (Adjusted Rand Index $ARI=0.869$). (B) UMAP embedding of the same data after randomisation of \emph{k}NN graph edges. In this case nSBM does not return any cell grouping, while optimisation of modularity finds up to 24 different cell groups}\label{FigureRandom}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[keepaspectratio,width=0.6\textwidth,height=\textheight]{FigureTian.png}
\caption[]{Analysis of a mixture of 5 known cell lines. (A) UMAP embedding coloured and labeled by cell line identity. H1975 cells (orange) show two distinct groups reflecting an internal heterogeneity. (B) UMAP embedding showing cells coloured by level 1 of the hierarchy proposed by the nested Stochastic Block Model. (C) UMAP embedding showing cells coloured according to the Leiden method at resolution $\gamma = 1$. (D) UMAP embedding coloured by the classification made by SCCAF when partitions in (B) are used. (E) UMAP embedding coloured by the classification made by SCCAF when partitions in (C) are used.}\label{FigureTian}
\end{figure}

@@

Identification of cells sharing similar properties in single cell experiments is of paramount importance. A large number of approaches have been described, although the standardisation of analysis pipelines converged to methods that are based on modularity optimisation. We tackled the biological problem using a different approach, nSBM, which has several advantages over existing techniques. The most important advantage is the hierarchical definition of cell groups which eliminates the choice of an arbitrary threshold on clustering resolution. In addition, we showed that the hierarchy itself could have a biological interpretation, implying that the hierarchical model is a valid representation of the cell ensemble.

>>>>

\begin{figure}[H]
\centering
\includegraphics[keepaspectratio,width=0.6\textwidth,height=\textheight]{FIgure_Random.png}
\caption[]{Modeling randomised data. (A) UMAP embedding of ~3k PBMC after standard processing with Leiden approach (left) or nSBM (right). Cell grouping is consistent for both the approaches (Adjusted Rand Index $ARI=0.869$). (B) UMAP embedding of the same data after randomisation of \emph{k}NN graph edges. In this case nSBM does not return any cell grouping, while optimisation of modularity finds up to 24 different cell groups}\label{FigureRandom}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[keepaspectratio,width=0.6\textwidth,height=\textheight]{FigureTian.pdf}
\caption[]{Analysis of a mixture of 5 known cell lines. (A) UMAP embedding coloured and labeled by cell line identity. H1975 cells (orange) show two distinct groups reflecting an internal heterogeneity. (B) UMAP embedding showing cells coloured by level 1 of the hierarchy proposed by the nested Stochastic Block Model. (C) UMAP embedding showing cells coloured according to the Leiden method at resolution $\gamma = 1$. (D) UMAP embedding coloured by the classification made by SCCAF when partitions in (B) are used. (E) UMAP embedding coloured by the classification made by SCCAF when partitions in (C) are used.}\label{FigureTian}
\end{figure}

@@

Identification of cells sharing similar properties in single cell experiments is of paramount importance. A large number of approaches have been described, although the standardisation of analysis pipelines converged to methods that are based on modularity optimisation. We tackled the biological problem using a different approach, nSBM, which has several advantages over existing techniques. \textcolor{red}{As  random data may have modular structure \cite{guimer_2004}, an important property of our approach is that it does not overfit data by finding partitions when, in fact, the are not.} Another important advantage is \textcolor{red}{that} the hierarchical definition of cell groups eliminates the choice of an arbitrary threshold on clustering resolution. In addition, we showed that the hierarchy itself could have a biological interpretation, implying that the hierarchical model is a valid representation of the cell ensemble.
